### 2018-03-16 paper/1-intro.tex
Implement some Simon's changes

Recently, the challenges of scale have driven large software firms
like Microsoft, Facebook, and Google to develop their own build
systems, each with its own choices and idiosyncrasies.
In this paper we offer a general framework in which to understand
and compare build systems,
in a way that is both abstract (omitting incidental detail)
and yet precise (implemented as Haskell code).  Specifically we make
these contributions:
\begin{itemize}
\item Build systems vary on many axes,
including: static vs dynamic dependencies; cloud-build, including
shallow vs deep; deterministic vs non-deterministic build rules;
early cut-off; self-tracking build systems; and persistent metadata.
In \S\ref{sec-background} we identify some key properties, in the context of
four carefully-chosen build systems.
\item We describe some simple but novel abstractions that
  crisply encapsulate what a build system is (\S\ref{sec-abstractions}),
  allowing us, for example, to speak about what it means for a build system to be correct.

>>>>

Recently, the challenges of scale have driven large software firms
like Microsoft, Facebook, and Google to develop their own build
systems, each with its own choices and idiosyncrasies.
In this paper we offer a general framework in which to understand
and compare build systems,
in a way that is both abstract (omitting incidental detail)
and yet precise (implemented as Haskell code).  Specifically we make
these contributions:
\begin{itemize}
\item Build systems vary on many axes, including:
static vs dynamic dependencies; local vs cloud; deterministic vs
non-deterministic build rules; support for early cutoff; self-tracking build
systems; and the type of persistent build information.
In \S\ref{sec-background} we identify some key properties, in the context of
four carefully-chosen build systems.
\item We describe some simple but novel abstractions that
  crisply encapsulate what a build system is (\S\ref{sec-abstractions}),
  allowing us, for example, to speak about what it means for a build system to be correct.

### 2018-03-16 paper/3-abstractions.tex
Implement some Simon's changes

\subsection{Common vocabulary for build systems}\label{sec-vocabulary}
\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- An abstract store
data Store i k v
getInfo    :: Store i k v -> i
putInfo    :: Store i k v -> i -> Store i k v
getValue   :: Store i k v -> k -> v
putValue   :: Eq k => Store i k v -> k -> v -> Store i k v
getHash    :: Hashable v => Store i k v -> k -> Hash v
initialise :: i -> (k -> v) -> Store i k v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Hashing
hash :: Hashable a => a -> Hash a
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Applicative functors
pure  :: Applicative f => a -> f a
(<$>) :: Applicative f =>   (a -> b) -> f a -> f b -- Left-associative
(<*>) :: Applicative f => f (a -> b) -> f a -> f b -- Left-associative
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard State monad from Control.Monad.State
data State s a
instance Applicative (State s)
instance Monad       (State s)
get       :: State s s
put       :: s -> State s ()
modify    :: (s -> s) -> State s ()
execState :: State s a -> s -> s
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Data.Functor.Identity and Data.Functor.Const
newtype Identity a = Identity { runIdentity :: a }
newtype Const m a  = Const    { getConst    :: m }
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty
    Const x <*> Const y = Const (x <> y)
\end{minted}
\vspace{-3mm}
\caption{Signatures of main data types and library functions.}\label{fig-types}
\vspace{-4mm}
\end{figure}
We begin by establishing a common vocabulary for build systems.
% that allows us to abstract away from specific application domains, such as
% software development or spreadsheets.
Fig.~\ref{fig-types} provides a Haskell encoding of our main definitions,
and gives types of the library functions we use in this paper.

\vspace{1mm}
A \emph{build system} takes a \emph{task description}, a target \emph{key},
and a \emph{store}, and returns a new store in which the target key is
up to date. The store is a key/value map, together with \emph{persistent build
information} that is maintained by the build system itself.

\vspace{-2mm}
\paragraph{Keys and values.} In software build systems, keys are typically filenames,
e.g.~\cmd{main.c}, and values are file contents (a C program source
code in this case). In spreadsheets, keys are cell names,
e.g. \cmd{A1}, and values are numbers, text, etc. that are typically
displayed inside cells. In our Haskell code, we will use type
variables \hs{k} and \hs{v} to denote keys and values, respectively.

@@

% \vspace{-2mm}
% \paragraph{The store.}
A \emph{store}, of type \hs{Store}~\hs{i}~\hs{k}~\hs{v}, associates keys to
values. It is convenient to assume that a store is total, i.e. it contains a
value for every possible key. We therefore also assume that the type of values
is capable of encoding values corresponding to non-existent keys (missing files
or empty cells). The store also holds a value of type \hs{i},
representing the persistent build information maintained by the build
system~--~its ``memory''.
% Given an \hs{i} and a key/value map \store, we can \hs{initialise} a store.

\vspace{-2mm}
\paragraph{Hashes.}
We use a cryptographic \emph{hash function}
\hs{hash}~\hs{::}~\hs{Hashable}~\hs{v}~\hs{=>}~\hs{v}~\hs{->}~\hs{Hash}~\hs{v}
for efficient tracking and sharing of build results, where \hs{Hash}~\hs{v} is
an abstract data type equipped with~the~equality test.
In addition to usual \hs{getValue} and \hs{putValue} operations,
some stores support a dedicated \hs{getHash} operation, which is implemented
more efficiently than by hashing the result of \hs{getValue}\footnote{For
example, Git Virtual File System~\cite{gvfs} stores file hashes and downloads
file contents only on demand.}.

% This, in particular, suggests that it is possible to \emph{depend on the
% absence of a value}, which is a useful feature for build systems.

>>>>

\subsection{Common vocabulary for build systems}\label{sec-vocabulary}
\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- An abstract store
data Store i k v
getInfo    :: Store i k v -> i
putInfo    :: Store i k v -> i -> Store i k v
getValue   :: Store i k v -> k -> v
putValue   :: Eq k => Store i k v -> k -> v -> Store i k v
getHash    :: Hashable v => Store i k v -> k -> Hash v
initialise :: i -> (k -> v) -> Store i k v
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Hashing
hash :: Hashable a => a -> Hash a
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Applicative functors
pure  :: Applicative f => a -> f a
(<$>) :: Applicative f =>   (a -> b) -> f a -> f b -- Left-associative
(<*>) :: Applicative f => f (a -> b) -> f a -> f b -- Left-associative
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard State monad from Control.Monad.State
data State s a
instance Applicative (State s)
instance Monad       (State s)
get       :: State s s
put       :: s -> State s ()
modify    :: (s -> s) -> State s ()
execState :: State s a -> s -> s
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard types from Data.Functor.Identity and Data.Functor.Const
newtype Identity a = Identity { runIdentity :: a }
newtype Const m a  = Const    { getConst    :: m }
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Functor (Const m) where
    fmap _ (Const m) = Const m
\end{minted}
\vspace{0.5mm}
\begin{minted}[fontsize=\small]{haskell}
instance Monoid m => Applicative (Const m) where
    pure _              = Const mempty
    Const x <*> Const y = Const (x <> y)
\end{minted}
\vspace{-3mm}
\caption{Signatures of main data types and library functions.}\label{fig-types}
\vspace{-4mm}
\end{figure}
We  begin by establishing a common vocabulary for build systems.

\vspace{-2mm}
\paragraph{Keys, values, and the store} The goal of any build system is to
bring up to date a \emph{store} that implements a mapping from \emph{keys} to
\emph{values}. In software build systems the store is the file system, the
keys are filenames, and the values are file contents. In \Excel, the store is
the worksheets, the keys are cell names (such as \cmd{A2}) and the values are
numbers, strings etc, displayed as the cell contents. Many systems use
\emph{hashes} of values as compact summaries with a fast equality check.

@@

\vspace{-2mm}
\paragraph{Persistent build information} As well as the key/value mapping, the store
also contains information maintained by the build system itself, which persists
from one invocation of the build system to the next -- its ``memory''.

\paragraph{Task description} Any build system requires the user to specify how to
compute the new value for one key, using the (up to date) values of its
dependencies. We call this specification the \emph{task description}.  For
example, in \Excel, the formulae of the spreadsheet constitute the task
description; in \Make the rules in the makefile are the task description.

\vspace{-2mm}
\paragraph{Build system} A \emph{build system} takes a task description, a target
key, and a store, and returns a new store in which the target key has an
up to date value.

\vspace{-2mm}
\paragraph{Modelling in Haskell} We will model all our build systems concretely,
as Haskell programs. To that end, Fig.~\ref{fig-types} gives the type
declarations and function signatures of the library functions. For example,
\hs{Store}~\hs{i}~\hs{k}~\hs{v} is the type of stores, with several associate
functions (\hs{getValue}, etc.). We use \hs{k} as a type
variable ranging over keys, \hs{v} for values, and \hs{i} for the persistent
build information.
